Tracing the Stellar Halos of Massive Galaxies to 100 kpc at $0.3<z<0.5$ using Hyper Suprime-Cam

### 2017-07-01 draft/paper1_v12.tex
Check Hirata2007
Small edits

% Version 0; preprint format; Written by SH
% Version 1; emulateapj format; Edited by SH and AL
% Version 2; emulateapj format; Edited by SH and AL
% Version 3; emulateapj format; Edited AL
% Version 4; emulateapj format; Edited SH
% Version 5; MNRAS format; Edited SH
% Version 6; MNRAS format; Paper I edited by JG
% Version 7; MNRAS format; Edited by SH
% Version 10; MNRAS format; Edited by AL and JG
% Version 11; MNRAS format; Edited by SH

@@

    %The mass-dependent nature we find in the \mden{} and ellipticity profiles of 
    %massive galaxies are consistent with the two-phase formation scenario. 
    We find that the outer halos of massive elliptical galaxies grow more prominent 
    and more elliptical with increasing stellar mass.
    According to the two-phase formation scenario, the inner 5--10 kpc of these 
    massive central galaxies are formed at $z>1$ during an intense period of 
    \textit{in situ} star formation. 
    The outskirts of massive galaxies are then built up through a more gradual second 
    phase of evolution (the \textit{ex situ} phase) that is dominated by mass assembly via 
    accretion via mergers.
    Non-dissipative mergers, especially the minor ones, mostly deposit stars in the 
    outskirts of centrals and do not have a large impact on the central \mden{} 
    profile (e.g. \citealt{BoylanKolchin2008, Oogi2013, Bedorf2013}).
    Given the stochastic nature of the merging process, it is easy to understand why 
    the \mden{} profiles of massive galaxies are similar in the inner region but show
    larger scatter in the outskirt.

@@

    Also according to this picture, the shape of the stellar halo should preserve
    information about the merging history and possibly even about the shape of the 
    dark matter halo. 
    Simulations show the shape of the stellar and dark matter halos are closely 
    correlated (e.g. \citealt{Wu2014}) for slowly-rotating massive ETGs having 
    undergone multiple minor-mergers.
    The more elongated outer halo and the trend between the ellipticity profile and 
    stellar mass may reflect the orbital properties of accreted satellite. 
    In simulations, satellite orbits become more radial for more massive halos 
    (e.g. \citealt{Murante2007, Wetzel2011, Jiang2015}. 
    Satellites infalling along radial orbits may help to form elongated stellar 
    halos. 
    Hence, in principle, these results help explain the trend that we see. 
    However, in contrast, simulation from \citet{Wu2014} predict that more massive 
    galaxies should have \emph{rounder} outskirts which is the opposite trend 
    compared to HSC. 
    This difference warrants further investigation.

    At larger scales, the distribution of satellites galaxies in massive halos is 
    found to be aligned with the major axis of the central galaxy 
    (e.g. \citealt{Brainerd2005, Yang2006b, NiedersteOstholt2010, 
    HuangMandelbaum2016}). 
    This alignment signal is found to be stronger for more massive galaxies 
    (e.g. \citealt{Hirata2007}).
    The fact that more massive central galaxies have steeper ellipticity profiles 
    and become increasingly more elongated in the outskirt may arise because of 
    such alignment. 
    Moreover, the shape of the dark matter halo can be estimated by analyzing 
    satellite distributions and weak lensing profiles (\citealt{ClampittJain2016}). 
    It is interesting to point out that the most recent measurement by 
    \citet{Shin2017} around SDSS clusters show a halo axis-ratio of ${\sim}0.55$, 
    which is only slightly more elongated than the stellar halo of galaxies in our 
    highest \mtot{} bin.

@@

    For more massive ETGs, results so far mainly depend on multi-component model 
    fitting and image stacking analyses (e.g. \citealt{Huang2013a, DSouza2014,
    Spavone2017}).
    In this work, we propose that \mstar{} computed within different fixed physical
    apertures are worth exploring as proxies of the \textit{ex situ} fraction. 
    We propose to use the mass within 10 kpc (\minn{}) and 100 kpc (\mtot{}) as 
    proxies for the \textit{in situ} component and for the total \mstar{}.

    On the left panel of Fig \ref{fig:discussion_1} we show the relation between 
    \mtot{} and the fraction of mass between 10 and 100 kpc (as a proxy of the mass of 
    the accreted component).
    We compare this with the fraction of \textit{ex-situ} stars predicted by the 
    Illustris simulation (\citealt{RodriguezGomez2016}).  
    We find that our proxy for the accreted mass component correlates strongly with  
    \mtot{} and that this relation is surprisingly consistent with predictions from
    \citet{RodriguezGomez2016}.  
    Given the limitations of the Illustris simulation, and the imperfect nature of 
    our \textit{ex situ} fraction proxy, the almost perfect agreement seen in 
    Figure \ref{fig:discussion_1} may well be a coincidence.  
    However, it is encouraging to see that our simple proxy for the fraction of 
    \textit{ex situ} stars seems to match both the slope and the scatter of the 
    predicted relation.  
    There exists other hydrodynamic simulations that predict significantly different 
    \textit{ex situ} fractions compared to Illustris (e.g. \citealt{Lackner2012, 
    Qu2017}). 
    In future work, we will explore more detailed comparisons between our data and
    predictions from hydrodynamic simulations and we will investigate to what degree 
    simple aperture masses may be used to trace \textit{ex situ} fractions.

@@

    We also compare with a uniquely massive BCG at high redshift: 
    a ${\sim} 10^{11.4} M_{\odot}$ BCG with a distinctive ``cD''-like envelope at 
    $z{\sim} 1.1$ (\citealt{Liu2013}).  
    This high redshift galaxy has a \mden{} profile that follows the median \mden{} 
    profile of our $11.6\leq$\logmtot{}$<11.8$ sample nicely at $\mathrm{R} < 20$ 
    kpc, but the profile becomes much steeper in the outskirt. 
    This suggests that the inner ``core'' of massive BCGs are already in pace at 
    $z{\sim} 1$ while the outer halo is still being assembled.

@@

        \item We find that the ``total'' \mstar{} of these massive galaxies can be 
            significantly underestimated with shallow imaging data such as SDSS and/or 
            oversimplified model assumptions (e.g. the \texttt{cModel} or single-\ser{}). 
            In contrast to previous work, our results do not depend on stacking or any 
            parametric models. 
            Moreover, the degree to which stellar mass is underestimated depends on 
            \mtot{}. 
            Simple model misses more light for massive galaxies because they have 
            more extended envelopes. 
            There is also a \mhalo{}-dependence of this effect and naive luminosity
            estimates will miss more light for BCGs in more massive halos compared 
            to centrals in less massive halos.
            These effects need to be carefully taken into account when discussing 
            topics such as the evolution of the galaxy stellar mass function.

@@

        \item We show that, on average, massive galaxies have positive ellipticity 
            gradients out to at least 60 kpc. 
            The average ellipticity profile also depend on \mtot{}: more massive 
            galaxies tend to have steeper ellipticity gradients and become more 
            elongated in the accreted stellar halo. 
            On the other hand, the average $(g-r)$ and $(g-i)$ color gradients do 
            not show clear dependence on \mtot{} within 10-60 kpc.

>>>>

% Version 0; preprint format; Written by SH
% Version 1; emulateapj format; Edited by SH and AL
% Version 2; emulateapj format; Edited by SH and AL
% Version 3; emulateapj format; Edited AL
% Version 4; emulateapj format; Edited SH
% Version 5; MNRAS format; Edited SH
% Version 6; MNRAS format; Paper I edited by JG
% Version 7; MNRAS format; Edited by SH
% Version 10; MNRAS format; Edited by AL and JG
% Version 11; MNRAS format; Edited by SH
% Version 12; MNRAS format; Edited by SH

@@

    %The mass-dependent nature we find in the \mden{} and ellipticity profiles of 
    %massive galaxies are consistent with the two-phase formation scenario. 
    We find that the outer halos of massive elliptical galaxies grow more prominent 
    and more elliptical with increasing stellar mass.
    According to the two-phase formation scenario, the inner 5--10 kpc of these 
    massive central galaxies are formed at $z>1$ during an intense period of 
    \textit{in situ} star formation. 
    The outskirts of massive galaxies are then built up through a more gradual second 
    phase of evolution (the \textit{ex situ} phase) that is dominated by mass assembly via 
    accretions.
    Non-dissipative mergers, especially minor mergers\footnote{Normally minor merger
    means the one with stellar mass ratio smaller than 1:3 or 1:4}, mostly deposit stars
    in the outskirts of centrals and do not have a large impact on the central \mden{} 
    profile (e.g. \citealt{BoylanKolchin2008, Oogi2013, Bedorf2013}).
    Given the stochastic nature of the merging process, it is easy to understand why 
    the \mden{} profiles of massive galaxies are similar in the inner region but show
    a large scatter in the outer region.

@@

    Also according to this picture, the shape of the stellar halo should preserve
    information about the merging history and possibly even about the shape of the 
    dark matter halo. 
    Simulations show the shape of the stellar and dark matter halos are closely 
    correlated (e.g. \citealt{Wu2014}) for slowly-rotating massive ETGs having 
    undergone multiple minor-mergers.
    The more elongated outer halo and the trend between the ellipticity profile and 
    stellar mass may reflect the orbital properties of accreted satellite. 
    In simulations, satellite orbits become more radial for more massive halos 
    (e.g. \citealt{Murante2007, Wetzel2011, Jiang2015}. 
    Satellites infalling along radial orbits may help to form elongated stellar 
    halos. 
    In principle, these results may explain the trend that we see. 
    However, in contrast, simulation from \citet{Wu2014} predict that more massive 
    galaxies should have \emph{rounder} outskirts which is the opposite trend 
    compared to HSC. 
    This difference warrants further investigation.

    At larger scales, the distribution of satellites galaxies in massive halos is 
    found to be aligned with the major axis of the central galaxy 
    (e.g. \citealt{Brainerd2005, Yang2006b, NiedersteOstholt2010, 
    HuangMandelbaum2016}). 
    This alignment signal is found to be stronger for more luminous galaxies 
    living in more massive halos (e.g. \citealt{Hirata2007}).
    The fact that more massive central galaxies have steeper ellipticity profiles 
    and become increasingly more elongated in the outskirt may arise because of 
    such alignment. 
    Moreover, the shape of the dark matter halo can be estimated by analyzing 
    satellite distributions and weak lensing profiles (\citealt{ClampittJain2016}). 
    It is interesting to point out that the most recent measurement by 
    \citet{Shin2017} around SDSS clusters show a halo axis-ratio of ${\sim}0.55$, 
    which is only slightly more elongated than the stellar halo of galaxies in our 
    highest \mtot{} bin.

@@

    For more massive ETGs, results so far mainly depend on multi-component model 
    fitting and image stacking analyses (e.g. \citealt{Huang2013a, DSouza2014,
    Spavone2017}).
    In this work, we propose that \mstar{} computed within different fixed physical
    elliptical apertures are worth exploring as proxies of the \textit{ex situ} 
    fraction. 
    We propose to use the mass within 10 kpc (\minn{}) and 100 kpc (\mtot{}) as 
    proxies for the \textit{in situ} component and for the total \mstar{}.

    On the left panel of Fig \ref{fig:discussion_1} we show the relation between 
    \mtot{} and the fraction of mass between 10 and 100 kpc (as a proxy of the mass of 
    the accreted component).
    We compare this with the fraction of \textit{ex-situ} stars predicted by the 
    Illustris simulation (\citealt{RodriguezGomez2016}).  
    We find that our proxy for the accreted mass component correlates strongly with  
    \mtot{} and that this relation is surprisingly consistent with predictions from
    \citet{RodriguezGomez2016}.  
    Given the limitations of the Illustris simulation, and the imperfect nature of 
    our \textit{ex situ} fraction proxy, the almost perfect agreement seen in 
    Figure \ref{fig:discussion_1} may well be a coincidence.  
    However, it is encouraging to see that our simple proxy for the fraction of 
    \textit{ex situ} stars seems to match both the slope and the scatter of the 
    predicted relation.  
    There exists other hydrodynamic simulations that predict significantly different 
    \textit{ex situ} fractions compared to Illustris (e.g. \citealt{Lackner2012, 
    Qu2017}). 
    In future work, we will explore more detailed comparisons between our data and
    predictions from hydrodynamic simulations and we will investigate to what degree 
    simple elliptical aperture masses may be used to trace \textit{ex situ} fractions.

@@

    We also compare with a uniquely massive BCG at high redshift: 
    a \logms{}${\sim} 10^{11.4} M_{\odot}$ BCG with a distinctive ``cD''-like envelope 
    at $z{\sim} 1.1$ (\citealt{Liu2013}).  
    This high redshift galaxy has a \mden{} profile that follows the median \mden{} 
    profile of our $11.6\leq$\logmtot{}$<11.8$ sample nicely at $\mathrm{R} < 20$ 
    kpc, but the profile becomes much steeper in the outskirt. 
    This suggests that the inner ``core'' of massive BCGs are already in pace at 
    $z{\sim} 1$ while the outer halo is still being assembled.

@@

        \item We find that the ``total'' \mstar{} of these massive galaxies can be 
            significantly underestimated with shallow imaging data such as SDSS and/or 
            oversimplified model assumptions (e.g. the \texttt{cModel} or single-\ser). 
            In contrast to previous work, our results do not depend on stacking or any 
            parametric models. 
            Moreover, the degree to which stellar mass is underestimated depends on 
            \mtot{}. 
            Simple model misses more light for massive galaxies because they have 
            more extended envelopes. 
            There is also a \mhalo{}-dependence of this effect and naive luminosity
            estimates will miss more light for BCGs in more massive halos compared 
            to centrals in less massive halos.
            These effects need to be carefully taken into account when discussing 
            topics such as the evolution of the galaxy stellar mass function.

@@

        \item We show that, on average, massive galaxies have positive ellipticity 
            gradients out to at least 60 kpc. 
            The average ellipticity profile also depend on \mtot{}: more massive 
            galaxies tend to have steeper ellipticity gradients and become more 
            elongated in stellar halos. 
            On the other hand, the average $(g-r)$ and $(g-i)$ color gradients do 
            not show clear dependence on \mtot{} within 10-60 kpc.
